Endogenous Politics and \\ the Design of Trade Institutions

### 2017-01-16 draft.tex
integrate explanation of \ga(e) into sec:economic

%\begin{assumption}
 % $\ga(s,e)$ is continuously differentiable, strictly increasing and concave in $e$.
  %\label{as:ga_c3}
%\end{assumption}

%Assumption~\ref{as:ga_c3} formalizes the intuition that the government favors the import-competing industry more the higher is its lobbying effort, but that there are diminishing returns to lobbying activity.

@@

We would like a framework in which the government's payoff is not constant. An appealing candidate is \Textcite{mrc2007}, which is in the spirit of \Textcite{gh94} but uses a bargaining framework. Although \Textcite{mrc2007} also assumes the constant returns form and that either the lobby or the government has all the bargaining power, \Textcite{lt} extend this framework to one with decreasing returns to lobbying and general bargaining weights.

@@

Depending on the level of $s$ and the properties of $\ga(s,e)$, the government may choose to set the trade agreement tariff in the presence of an escape clause $\left(\tau_e^R\right)$ so that in some cases the lobby's incentive constraint is violated. This would lead to the legislature's constraint being violated as well, creating a dispute.

>>>>

\begin{assumption}
  $\ga(s,e)$ is continuously differentiable, strictly increasing and concave in $e$.
  \label{as:ga_c3}
\end{assumption}

Assumption~\ref{as:ga_c3} represents the idea that the government favors the import-competing industry more the higher is its lobbying effort, but that there are diminishing returns to lobbying activity.

This objective function for the government is consistent with \Textcite{bm93}, which explores advertising and the goods being advertised as complementary in a `stable metautility function.' Their consumers are sophisticated and can advocate for policy to change the level of advertising with which they will be confronted. Similarly, the government here has a stable mapping from the amount of lobbying effort it experiences to its optimal choice of tariffs and it can use the trade agreement to change the level of lobbying effort with which it will be confronted. The government's preferences don't change with $e$, as the preferences related to lobbying effort are represented by the mapping $\ga(\cdot)$. When $e$ changes, the metautility function is evaluated at a different value.

As Figure~\ref{fig:ext} makes clear, lobbying, as well as the realization of any shocks, occurs after the trade agreement is negotiated and before the applied tariff decision. $\ga(s,e)$ is thus evaluated at the ex-post stage, while the trade agreement is negotiated in the shadow of the realizations of $s$ and $e$. We can thus interpret the fundamental structure of preferences in a couple of ways. We can think of lobbying effort as a kind of persuasion than affects decision making of a single actor or we can think of lobbying effort as influencing the identify of the ex-post decision-maker. In either case, $\ga(s,e)$ is the weight the decision-maker places on the lobbying industry's profits ex-post, and the realization of this weight changes with the realization of the shock and lobbying effort even while the mapping---the preferences---do not.\footnote{The model also admits an interpretation in which the decision-making body is non-unitary, such as a large legislature. Here, $s$ and $e$ determine the identity of the median legislator. This interpretation results in minor, qualitative changes to the repeated-game results because the non-unitary decision-maker knows that different decision-makers will be in power in the future and evaluates future welfare with her own preferences. In contrast, the unitary decision-maker evaluates future welfare according to the lobbying effort she expects to experience in the future.}

@@

We would like a framework in which the government's payoff is not constant. An appealing candidate is \Textcite{mrc2007}, which is in the spirit of \Textcite{gh94} but uses a bargaining framework. In my setup, the lobby is not active during the trade agreement negotiation; in order to to fit into the framework of \Textcite{mrc2007}, we can assume that capital is perfectly mobile in the long run so that it is not worthwhile for the lobby to expend resources to influence the negotiation of the trade agreement.

Although \Textcite{mrc2007} also assumes the constant returns form and that either the lobby or the government has all the bargaining power, \Textcite{lt} extend this framework to one with decreasing returns to lobbying and general bargaining weights.

@@

Depending on the level of $s$ and the properties of $\ga(s,e)$, the government may choose to set the trade agreement tariff in the presence of an escape clause $\left(\tau_e^R\right)$ so that in some cases the lobby's incentive constraint is violated. This would lead to the government's constraint being violated as well, creating a dispute.

### 2017-01-16 to-do.tex
integrate explanation of \ga(e) into sec:economic

\section{``Dynamic preferences''}
	\begin{itemize}
		\item Becker, G. S. and K. M. Murphy (1993), Simple Theory of Advertising as a Good or Bad, Quarterly Journal of Economics, 942-64.
			\begin{itemize}
				\item Their firms are my firm. Their consumers are my government
				\item Kyle's lit review talkes about `persuasive,' `informative,' and `complementary' views of advertisting. This is complementary.
				\item $e$ and $\tau$ enter objective function. Gov't is sophisticated, can make commitment that alters $e$ (i.e. behavior of another actor). But $CS + \ga(e)PS + \ldots$, just like $u(X,Y,A)$; says consumer could work to pass law about number of ads on TV each hour.
				\item Use newly drawn extensive form to explain this, lays out when welfare is realized
				\item Addresses James' concern about needing to be more precise about when $\ga$ is evaluated
					\begin{itemize}
						\item Skeptical of comparative-static style result: can I really compare governments with different $\ga$'s?
					\end{itemize}
			\end{itemize}
		\item Mostafa, Nuno, Giovanni, Anson to some extent: Cardinal vs. ordinal
				\begin{itemize}
					\item Giovanni: if I can show properties survive  monotonic transformation, I'm fine
					\item They don't like gov't choosing its own preferences. May just need to change wording
				\end{itemize}
		\item Bob: what is fundamental structure of preferences?
		\item The way $e$ affects preferences doesn't change!!! Preferences are wrapped up in $\ga$!
			\begin{itemize}
				\item ``How the assumption of a non-unitary legislature translates into this incentive constraint. In the non-unitary legislature, if the current period incentive constraint includes any future value of lobbying effort, this would mean the current-period median legislator evaluating her current-period incentive constraint using a mixture of her own political-economy weight with those of the legislators who are median in the future in the trade agreement and trade-war scenarios. The first major step I have taken to clarify the incentive constraint is the analysis in Appendix B.1 of the unitary model. I discuss the differences between the unitary and non-unitary models in the beginning of that Appendix.''
				\item ``Note that the median legislator, whose identity is determined by the lobby's effort level $e_b$, evaluates future payoffs according to her own political economy weight, $\ga(e_b)$. Of course, depending on legislator $e_b$'s choice, either legislator $e_a$ or legislator $e_{tw}$ will be the decision maker in those future periods. But legislator $e_b$, who is the decision maker in the current period, maximizes her own welfare given the predicted behavior of future decision makers.''
			\end{itemize}
		\item Devashish: Utility fcn didn't change. Just got evaluated at a different point
				\begin{itemize}
					\item Preferences don't change. It's persuasion that affects decision-making. 2nd story: Giovanni's ``expost $\ga(e)$ as effort to elect/choose ex-post decision-maker''
				\end{itemize}
								\item Unitary vs. non-unitary
					\begin{itemize}
						\item Assume the lobby is not active during the ex-ante phase, that is, when the trade agreement is being negotiated. At the ex-ante stage, $e=0$. If this government is also unitary, then this interpretation fits into the framework of Maggi and Rodriguez-Clare 2007 by assuming that capital is perfectly mobile in the long run so that it is not worthwhile for the lobby to expend resources to influence the negotiation of the trade agreement.
						\item Note that this remains a non-unitary model. Out of a single decision-making body, the level of lobbying effort determines a different decisive member depending on the situation, e.g. during ex-ante negotiations ($e=0$) versus a trade war ($e=e_{tw}$). A unitary model---in which a single actor makes different decisions depending on how much lobbying effort she experiences---results in minor, qualitative changes to the results. Note that in either case, it is only the realization of $\ga(\cdot)$ that changes with $e$, not the preferences themselves which are embodied in $\ga(\cdot)$.
						\item To be clear about the differences, the non-unitary model examined in the body of the paper assumes that different levels of lobbying effort result in different \textit{decision-makers}. In the unitary model of DGH97, one decision maker makes different decisions depending on the level of lobbying effort. This distinction is mainly important in the context of the repeated game. Whereas the unitary decision-maker evaluates future welfare according to the lobbying effort she expects to experience in the future, the non-unitary decision-maker knows that different decision-makers will be in power and evaluates future welfare with her own preferences.
						\item In the model with a non-unitary government, the lobby's effort $e$ determines the identity of the decision maker. Thus the decision maker at the time the lobby is pushing for the agreement to be broken ($e=e_b>e_a$ in order to have a median legislator who will break the agreement) is different from the decision maker during the trade war phase $(e = e_{tw})$ and the decision maker during a trade agreement phase ($e=e_a(\tau^a)$). The weight on the lobbying industry's profits changes with lobbying effort because lobbying effort determines a different median voter in the legislature or government more generally. However, when the median voter during a `break' phase is evaluating her incentive constraint, she values future tariff choices --- which will be determined by whoever is median in the future --- with her own preferences. Although she will not be the decision maker in the future, there is no reason for her to evaluate future welfare according to some other legislator's preference.
						\item If the model is interpreted as having a unitary government, there is only one decision maker. There is a fixed mapping from lobbying effort to the weight the decision maker places on the lobby's profits, and the realization of this weight changes with the realization of lobbying effort. Thus, predicting future lobbying effort, the decision maker knows what the realization of her weight on the lobby's profits will be and evaluates the incentive constraint accordingly. Expression (10) would be modified while the lobbying constraint is unchanged.
					\end{itemize}
		\end{itemize}

@@

\section{Misc.}
\begin{itemize}
	\item Specialize to BS2005 when necessary (tariff cap example)
	\item Define $\ga$ before using it
	\item If add back in assumption 1, add Ethier footnote
	\item standardize $\ga(e,s)$, $\ga(e)$, $\ga(s)$ throughout the text
	\item Peter: $\ga(e)$ and $\ga(s)$ combination is interesting, so it dynamic choice of protection over time, gov't turnover [add to conclusion, it's in slides]
	\item Ben: I'm wondering how you view the time-frame over which your model is set.  On one hand a specific-factors set-up is normally associated with the short-run.  But on the other hand you have an infinite time horizon which suggests very long run.
	\item Clarify information structure
	\item Move footnote about TOT externality (fn:tot) into text and make paragraphs flow together
	\item Add ref to JIE R$\&$R to fn:krw ?
	\item Expand Proposition 2 (res:repeated) to compare to exogenous case, say what $\tau^R_{W,e}$ IS. Perhaps not all of that IN the proposition.
	\item Need more sign-posting in the three last paragraphs of Section 4.2. Maybe make Section 4.3 a new top-level section.
\end{itemize}

>>>>

@@

\section{Misc.}
\begin{itemize}
	\item Specialize to BS2005 when necessary (tariff cap example)
	\item They don't like gov't choosing its own preferences. May just need to change wording.
	\item Define $\ga$ before using it
	\item If add back in assumption 1, add Ethier footnote
	\item standardize $\ga(e,s)$, $\ga(e)$, $\ga(s)$ throughout the text
	\item Peter: $\ga(e)$ and $\ga(s)$ combination is interesting, so it dynamic choice of protection over time, gov't turnover [add to conclusion, it's in slides]
	\item Ben: I'm wondering how you view the time-frame over which your model is set.  On one hand a specific-factors set-up is normally associated with the short-run.  But on the other hand you have an infinite time horizon which suggests very long run.
	\item Clarify information structure
	\item Move footnote about TOT externality (fn:tot) into text and make paragraphs flow together
	\item Add ref to JIE R$\&$R to fn:krw ?
	\item Expand Proposition 2 (res:repeated) to compare to exogenous case, say what $\tau^R_{W,e}$ IS. Perhaps not all of that IN the proposition.
	\item Need more sign-posting in the three last paragraphs of Section 4.2. Maybe make Section 4.3 a new top-level section.
\end{itemize}
